Learning to Dispatch: A Reinforcement Learning Framework for Train Dispatch Networks

### 2025-04-23 paper/main.tex
Some typos and business

% Paper Outline
To provide a thorough understanding of the approach, the remainder of this paper is organized as follows. 
In Section~\ref{sse:background}, we present the foundational background necessary for our formulation, beginning with a formal description of the Train Dispatch Problem, reinforcement learning, and graph neural networks.
These concepts are then reinforced in a discussion of recent research that inspired this work.
Section~\ref{sse:formulation} details the proposed formulation, introducing Train Operation Graphs (\ref{sss:train_ops}), Resource Conflict Graphs (\ref{sss:resource_conflicts}), and the state and action spaces (\ref{sss:state_space}, \ref{sss:action_space}). 
In Section~\ref{sse:discussion}, preliminary results are showcased from a graph-based Deep Q-Network (DQN) agent followed by a discussion of the limitations behind the current DQN implementation.
Finally, Section~\ref{sse:conclusion} concludes this work and discusses future directions for scaling and refining the framework.

@@

Devailly et al.~\cite{gnndrl:Devailly_2022} introduced IG-RL, an inductive graph reinforcement learning approach for large-scale traffic signal control. 
By utilizing GNNs, IG-RL efficiently manages massive traffic networks, enabling scalable and adaptive control strategies. 
Similarly, Agasucci et al.~\cite{gnndrl:agasucci} proposed a DRL-based formulation for the train dispatching problem, although their approach is constrained by the specificity of the data source used.

@@

\section{Formulation}
\label{sse:formulation}

@@

\section{Discussion}
\label{sse:discussion}
\subsection{Preliminary Agent Results}

\subsection{Implementation Limitations}

\section{Conclusion and Future Work}
\label{sse:conclusion}

@@

This honors thesis introduces a proof-of-concept reinforcement learning framework for the Train Dispatching Problem, emphasizing the formulation and potential of graph-based representations and deep learning techniques. 
By framing the problem as a Markov Decision Process and leveraging Graph Neural Networks, this work provides a foundational approach to train dispatching that is both scalable and extensible.

>>>>

% Paper Outline
To provide a thorough understanding of the approach, the remainder of this paper is organized as follows. 
In Section~\ref{sse:background}, we present the foundational background necessary for our formulation, beginning with a formal description of the Train Dispatch Problem, reinforcement learning, and graph neural networks.
These concepts are then reinforced in a discussion of recent research that inspired this work.
Section~\ref{sse:formulation} details the proposed formulation, introducing Train Operation Graphs (\ref{sss:train_ops}), Resource Conflict Graphs (\ref{sss:resource_conflicts}), and the state and action spaces (\ref{sss:state_space}, \ref{sss:action_space}). 
Finally, Section~\ref{sse:discussion} preliminary results are showcased from a graph-based Deep Q-Network (DQN) agent followed by future directions for work.

@@

Devailly et al.~\cite{gnndrl:Devailly_2022} introduced IG-RL, an inductive graph reinforcement learning approach for large-scale traffic signal control. 
By utilizing GNNs, IG-RL efficiently manages massive traffic networks, enabling scalable and adaptive control strategies. 
Similarly, Agasucci et al.~\cite{gnndrl:agasucci} introduced a DRL-based approach to the train dispatching problem, though their method is limited by the specificity of the dataset employed.

@@

\section{Formulation}
\label{sse:formulation}
The proposed formulation introduces a approach to the TDP by leveraging both graph-based representations and reinforcement learning techniques. 
By modeling train operations as Directed Acyclic Graphs (DAGs) and resource conflicts as Resource Conflict Graphs (RCGs), this formulate aims to capture the intricate dependencies and constraints inherent to railway systems. 
The integration of Graph Neural Networks (GNNs) enables efficient processing of these graph structures, while the reinforcement learning agent learns to make optimal dispatching decisions dynamically.

The remainder of this section is organized as follows. 
Subsection~\ref{sss:train_ops} introduces the concept of Train Operation Graphs in the DISPLIB format, detailing their structure and role in representing train operations and constraints. 
Subsection~\ref{sss:resource_conflicts} describes the Resource Conflict Graphs, which model resource-sharing conflicts between trains. Subsection~\ref{sss:state_space} defines the state space, highlighting how graph-based representations encode the environment's current state. 
Finally, Subsection~\ref{sss:action_space} outlines the action space, specifying the set of actions available to the reinforcement learning agent for decision-making.

@@

\section{Discussion and Future Work}
\label{sse:discussion}

\subsection{Discussion}
\label{sss:discussion}

@@

This honors thesis introduces a proof-of-concept reinforcement learning framework for the Train Dispatching Problem alongside preliminary results, emphasizing the formulation and potential of graph-based representations and deep learning techniques. 
By framing the problem as a Markov Decision Process and leveraging Graph Neural Networks, this work provides a foundational approach to train dispatching in the DISPLIB format that is both scalable and extensible.
